---
title: "Homework: Week 3 - Logistic Regression"
output:
  html_notebook:
    toc: yes
    toc_float:
      collapsed: true
      smooth_scroll: true
  github_document: default
---

# Popularity of Music Records

## Load the data 

```{r, message=FALSE}
library(readr)
library(tidyverse)
songs <- read_csv("~/Documents/Online Classes/analytics-edge/songs.csv")
```

##EDA

How many observations (songs) are from the year 2010?

```{r}
table(songs$year)
```

How many songs does the dataset include for which the artist name is "Michael Jackson"?

```{r}
songs %>% filter(artistname == "Michael Jackson") %>% count()
```

Which of these songs by Michael Jackson made it to the Top 10?

```{r}
(songs %>% filter(artistname == "Michael Jackson", Top10 == 1))
```

What are the values of `timesignature` that occur in our dataset?

```{r}
sort(unique(songs$timesignature))
```

Which `timesignature` value is the most frequent among songs in our dataset?

```{r}
table(songs$timesignature)
```

Which song in our dataset has the highest tempo?

```{r}
which.max(songs$tempo)
songs$songtitle[6206]

songs %>% filter(tempo == max(tempo))
```

##Model Making

How many observations (songs) are in the training set?

- Training set: $\leq 2009$
- Test set: $\geq 2010$

```{r}
SongsTrain <- subset(songs, songs$year <= 2009)
SongsTest <- subset(songs, songs$year >= 2010)
nrow(SongsTrain)
```

Exclude non-integer variables:

```{r}
nonvars = c("year", "songtitle", "artistname", "songID", "artistID")

SongsTrain = SongsTrain[ , !(names(SongsTrain) %in% nonvars) ]
SongsTest = SongsTest[ , !(names(SongsTest) %in% nonvars) ]
```

Build a model w/ training set:

```{r}
SongsLog1 = glm(Top10 ~ ., data=SongsTrain, family="binomial")
summary(SongsLog1)
```

## Beware Collinearity


What is the correlation between the variables "loudness" and "energy" in the training set?

```{r}
cor(SongsTrain$loudness, SongsTrain$energy)
```

Create a second model, which excludes "loudness":

```{r}
SongsLog2 = glm(Top10 ~ . - loudness, data=SongsTrain, family="binomial")
summary(SongsLog2)
```

Create Model 3, which is "Model 1" without "energy"

```{r}
SongsLog3 = glm(Top10 ~ . - energy, data=SongsTrain, family=binomial)
summary(SongsLog3)
```

## Test the Model

We will be using Model 3 to validate the model

```{r}
PredictTest <- predict(SongsLog3, newdata = SongsTest, type = "response")

table(SongsTest$Top10, PredictTest > 0.45)

(309 + 19)/(nrow(SongsTest))
```

What would the accuracy of the "baseline"" model be on the test set?

```{r}
table(SongsTest$Top10)
314/nrow(SongsTest)
```

How many songs does Model 3 correctly predict as Top 10 hits in 2010 (threshold 0.45)?

```{r}
table(SongsTest$Top10, PredictTest > 0.45)
```

What is the sensitivity of Model 3 on the test set, using a threshold of 0.45? specificity?

```{r}
library(pROC)
ROC_test <- roc(SongsTest$Top10, PredictTest)
coords(ROC_test, 0.45, ret = c("sens", "spec"))
```

# Predicting Parole Violators

## Load the Data

```{r, message=FALSE}
parole <- read_csv("~/Documents/Online Classes/analytics-edge/parole.csv")
glimpse(parole)
```

## EDA

How many of the parolees in the dataset violated the terms of their parole?

```{r}
table(parole$violator)
```

Convert state, crime to unordered **factors**?

```{r}
parole$state <- as.factor(parole$state)
parole$crime <- as.factor(parole$crime)
summary(parole)
```

## Create Train vs Test

```{r}
set.seed(144)

library(caTools)

split = sample.split(parole$violator, SplitRatio = 0.7)

train = subset(parole, split == TRUE)
test = subset(parole, split == FALSE)
```

## Build a Model

```{r}
parole_model <- glm(violator ~ ., data = train, family = "binomial")
summary(parole_model)
```

## Test the Model

What is the maximum predicted probability of a violation?

```{r}
predict_parole <- predict(parole_model, newdata = test, type = "response")

max(predict_parole)
```

What is the model's sensitivity? specificity? accuracy?

```{r}
ROC_parole <- roc(test$violator, predict_parole)
coords(ROC_parole, 0.5, ret = c("sens", "spec", "acc"))
```

What is the accuracy of a simple model that predicts that every parolee is a non-violator?

```{r}
table(test$violator)

179/nrow(test)
```

What is the AUC value for the model?

```{r}
auc(ROC_parole)
```

# Predicting Loan Repayment

## Load the Data

```{r, message=FALSE}
loans <- read_csv("~/Documents/Online Classes/analytics-edge/loans.csv")
glimpse(loans)
```

What proportion of the loans in the dataset were not paid in full?

```{r}
table(loans$not.fully.paid)

1533/nrow(loans)
```

Which of the following variables has at least one missing observation? 

```{r}
summary(loans)
```

